---
title: "p8105_hw3_yw3996"
author: "Yiying Wu"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## 1. Instacart data
load the data from the p8105.datasets using:
```{r,include=TRUE,results = 'hide',warning = FALSE, message = FALSE}
library(p8105.datasets)
data("instacart")
instacart=instacart
```
### Description of the dataset

* This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. 
* Variables include 
  
  * `order_id`: order identifier
  * `product_id`: product identifier
  * `add_to_cart_order`: order in which each product was added to cart
  * `reordered`: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
  * `user_id`: customer identifier
  * `eval_set`: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train” eval_set)
  * `order_number`: the order sequence number for this user (1=first, n=nth)
  * `order_dow`: the day of the week on which the order was placed
  * `order_hour_of_day`: the hour of the day on which the order was placed
  * `days_since_prior_order`: days since the last order, capped at 30, NA if order_number=1
  * `product_name`: name of the product
  * `aisle_id`: aisle identifier
  * `department_id`: department identifier
  * `aisle`: the name of the aisle
  * `department`: the name of the department

### Count the number of aisles and the aisles with most items ordered
Count the number of aisles in the dataset:
```{r}
n_distinct(instacart$aisle)
```

Count the number of orders in each aisle and sort by descending order:
```{r}
instacart |>
  group_by(aisle) |>
  summarize(n_obs = n())|>
  arrange(desc(n_obs))
```
Therefore, there are `r n_distinct(instacart$aisle)` aisles. The most items are ordered from fresh vegetables aisles.

### Plot 1 
the number of items ordered in each aisle (aisles with more than 10000 items ordered)

```{r}
instacart |>
  group_by(aisle) |>
  summarize(n_obs = n()) |>
  filter(n_obs > 10000) |>
  mutate(aisle = fct_reorder(aisle, desc(n_obs))) |>
  ggplot(aes(x = aisle, y = n_obs)) + 
  geom_bar(stat = "identity") +  # Use geom_bar for a bar chart
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
Comments:

The bar chart depicted above illustrates the quantity of items requested from various store aisles, specifically focusing on those aisles with an order count exceeding 10,000 items. The aisles have been organized in a descending order according to their respective item counts, simplifying the identification of the most frequented sections. The fresh vegetables aisle emerges as the most popular, boasting the highest item order count, followed by the fresh fruits and packaged vegetables fruits sections. Conversely, aisles like oils vinegars, dry pasta, canned meals beans, and butter exhibit the lowest order counts among the ones presented. Nonetheless, even these less popular sections still manage to accumulate more than 10,000 orders.

### Table 1 
the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” with the number of times each item is ordered
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", 
                      "dog food care", 
                      "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```
Comments:

This table provides information on the top-ranked products within various categories in a store. 

* Packaged Vegetables & Fruits:

  * Organic Baby Spinach: This product is the most popular in the "Packaged Vegetables Fruits" category, with a substantial count of 9784. It's evident that customers prefer healthy, organic greens.
  * Organic Raspberries: Ranked second, "Organic Raspberries" also enjoys significant popularity with 5546 purchases.
  * Organic Blueberries: "Organic Blueberries" holds the third position, highlighting the appeal of organic fruits among shoppers.

* Baking Ingredients:

  * Light Brown Sugar: In the "Baking Ingredients" category, "Light Brown Sugar" tops the list with 499 purchases. It's a staple for baking enthusiasts.
  * Pure Baking Soda: "Pure Baking Soda" is the second most purchased product within this category, reflecting its importance in baking.
  * Cane Sugar: While "Cane Sugar" ranks third, it still enjoys a notable 336 purchases, indicating the demand for various sweeteners in baking.

* Dog Food Care:

  * Snack Sticks Chicken & Rice Recipe Dog Treats: This product is the most popular in the "Dog Food Care" category, showing that pet owners prioritize treats for their dogs.
  * Organix Chicken & Brown Rice Recipe: Ranked second, "Organix Chicken & Brown Rice Recipe" caters to health-conscious pet owners.
  * Small Dog Biscuits: This product is the third most purchased, demonstrating that even small dog breeds have their specific dietary needs.

### Table 2 
the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day,na.rm=TRUE),.groups="drop") |>
  mutate(order_dow = case_when(
    order_dow == 0 ~ "Monday",
    order_dow == 1 ~ "Tuesday",
    order_dow == 2 ~ "Wednesday",
    order_dow == 3 ~ "Thursday",
    order_dow == 4 ~ "Friday",
    order_dow == 5 ~ "Saturday",
    order_dow == 6 ~ "Sunday"
  ))|>
  pivot_wider(
    names_from = "order_dow", 
    values_from = "mean_hour")|>
  knitr::kable(digits = 2)
```

Comments:

The table presents the typical ordering times for "Coffee Ice Cream" and "Pink Lady Apples" throughout the week. In the case of "Coffee Ice Cream," the mean order time falls within the early to mid-afternoon range for all days. It commences at approximately 1:45 PM on Monday and reaches its latest around 3:30 PM on Thursday. Conversely, "Pink Lady Apples" consistently exhibit an average ordering time in the early afternoon for the entire week. This timing begins at roughly 11:30 AM on Tuesday and experiences slight variations but generally remains within the early to mid-afternoon timeframe for the remaining days.

## 2. BRFSS data
load the data from the p8105.datasets using:
```{r,include=TRUE,results = 'hide',warning = FALSE, message = FALSE}
data("BRFSS")
brfss=brfss_smart2010
```

### data cleaning

* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Poor” to “Excellent”

list all the value of response column:
```{r}
brfss=brfss|>
  janitor::clean_names()|>
  mutate(data_value=as.double(data_value))
unique(brfss$response)
```
The responses should be included are "Excellent", "Very good",  "Good", "Fair", "Poor".

The order should be "Poor", "Fair", "Good", "Very good", "Excellent"

```{r}
brfss=brfss|>
  filter(
    topic=="Overall Health",
    response %in% c("Poor", "Fair", "Good", 
                    "Very good", "Excellent"))|>
  mutate(
    response=factor(response, levels=c("Poor", "Fair", "Good", 
                                       "Very good", "Excellent"))
  )
```

### states observed at 7 or more locations in 2002 or in 2010 

states were observed at 7 or more locations in 2002 are: 
```{r}
brfss|>
  filter(year==2002)|>
  group_by(locationabbr)|>
  summarize(n_obs = n_distinct(locationdesc))|>
  filter(n_obs>=7)
```
CT, FL, MA, NC, NJ, PA were observed at 7 or more locations in 2002.

states were observed at 7 or more locations in 2010 are:
```{r}
brfss|>
  filter(year==2010)|>
  group_by(locationabbr)|>
  summarize(n_obs = n_distinct(locationdesc))|>
  filter(n_obs>=7)
```

CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA were observed at 7 or more locations in 2010.

### Construct a dataset with Excellent responses including year, state, and a variable that averages the data_value across locations within a state

```{r}
dat = brfss|>
  filter(response=="Excellent")|>
  group_by(year, locationabbr)|>
  summarize(data_value_mean=mean(data_value, na.rm = TRUE))
```
### spaghetti plot
```{r, fig.width = 8}
dat|>
  rename("state"="locationabbr")|>
  ggplot(aes(x = year, y = data_value_mean, 
             color =state )) + 
  geom_point() + geom_line() + 
  theme(legend.position = "bottom")+
  guides(color = guide_legend(nrow = 5))
```
Comments:

The average data_value for excellent response fluctuate from 2002 to 2010. For most of the states average data_value for excellent response is roughly between 17 to 27.

### two-panel plot
two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r, fig.width = 9}
plot=
  brfss |>
  filter(locationabbr=="NY"& year %in% c(2006,2010))|>
  ggplot(aes(x =response , y = data_value)) + 
  geom_boxplot()+
  labs(title = "Distribution of data_value in NY State in 2006 and 2010") +
  facet_grid(.~year)

plot
```
Comments:

* In both 2006 and 2010, the "Poor" responses exhibit the lowest median data values, and the "Good ", "Very Good", and "Excellent" responses tend to have significantly higher data values compared to those with "Poor" or "Fair" responses.
* The "Fair" responses in 2010 have a greater variation compared to those in 2002.
* The "Good" responses in 2010 have smaller variation compared to those in 2002.
* The "Very Good" responses in 2010 have a higher median and larger variation compared to those in 2002.

two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) in different  locations in NY State.

```{r}
plot_2006=
  brfss |>
  filter(locationabbr=="NY",year==2006)|>
  ggplot(aes(x =locationdesc , y = data_value,fill=response)) + 
  geom_bar(position="dodge",stat="identity") +  
  labs(title = "Distribution of data_value in NY State in 2006") + 
  theme(plot.title = element_text(size = 10), axis.text.x = element_text(angle = 20, hjust = 1),legend.position = "none") 

plot_2010=
  brfss |>
  filter(locationabbr=="NY",year==2010)|>
  ggplot(aes(x =locationdesc , y = data_value,fill=response)) + 
  geom_bar(position="dodge",stat="identity") +  
  labs(title = "Distribution of data_value in NY State in 2010") +
  theme(plot.title = element_text(size = 10), axis.text.x = element_text(angle = 20, hjust = 1))

plot_2006 + plot_2010
```
Comments:

There are 6 counties responded in 2006, while 9 counties in 2010. The overall distribution seems consistent between two years, while there is variation among counties.

## 3. Accelerometer data
import dataset
```{r, include=TRUE,results = 'hide',warning = FALSE, message = FALSE}
demographic=read_csv("./data/nhanes_covar.csv",skip = 4)|>
  janitor::clean_names()
accelerometer=read_csv("./data/nhanes_accel.csv")|>
  janitor::clean_names()
```

### data cleaning
clean `demographic` dataset:

* include all originally observed variables; 
* exclude participants less than 21 years of age, and those with missing demographic data; and 
* encode data with reasonable variable classes
```{r}
demographic=demographic|>
  drop_na()|>
  filter(age>=21,!is.na(sex), !is.na(age), !is.na(bmi), !is.na(education))|>
  mutate(
     sex=case_match(
       sex,
       1~"male",
       2~"female"),
     education=case_match(
       education,
       1~"Less than high school",
       2~"High school equivalent",
       3~"More than high school"
     ),
     education=factor(education, levels=c(
       "Less than high school",
       "High school equivalent",
       "More than high school"))
  )
```

clean `accelerometer` data

* pivot longer
```{r}
accelerometer=accelerometer|>
  drop_na()|>
  pivot_longer(
    min1:min1440,
    names_to = "time", 
    values_to = "mims_value")|>
  separate(time,into=c("min","minute"),3)|>
  select(-min)|>
  mutate(minute=as.integer(minute))
```

join `accelerometer` with `demographic`  
```{r}
dat_3=
  inner_join(demographic,accelerometer,by="seqn")
```

### table for the number of men and women in each education category
```{r}
dat_3|>
  group_by(sex,education) |> 
  summarize(n_obs = n_distinct(seqn)) |> 
  pivot_wider(names_from = education, values_from = n_obs)|>
  knitr::kable()
```
Comments: 

* Among both genders, the "More than high school" education category has the highest count, indicating that a significant portion of the sampled population has education beyond high school.
* In the "High school equivalent" category, there are slightly more males (35) than females (23).
* The counts in "Less than high school" category are close, with 28 females and 27 males. 

### visualization of the age distributions for men and women in education category
```{r}
dat_3 |> 
  group_by(education,seqn)|>
  distinct(seqn,sex,age,education)|>
  ggplot(aes(x = age, fill = education)) + 
  labs(title = "Male") +
  geom_density(alpha = .5) +
  facet_grid(.~sex)
```
Comments: 

* Among both genders, the age distributions for "More than high school" education category are right-skewed, while age distributions for "Less than high school" education category are left-skewed, indicating that young people tend to have higher education levels.
* The difference of the distributions among education categories for female is more significant than that of male group. This suggests that gender may play a role in shaping educational attainment patterns, with females showing a greater variation in age across education levels.


### analysis total activity variable for each participant
create a total activity variable for each participant
```{r}
total_act=dat_3|>
  group_by(seqn,sex,age,bmi,education)|>
  summarise(total_activity=sum(mims_value))
```
Plot these total activities (y-axis) against age (x-axis)
```{r}
total_act|>
  ggplot(aes(x = age, y = total_activity, color = sex)) + 
  geom_point(alpha = .75)+
  geom_smooth()+
  facet_grid(.~education)
```

Comments:

* Overall, both young females and young males tend to have higher total activities in all education groups.
* For "High school equivalent" and "More than high school" groups, the total activities for females are slightly higher than those for males.
* Since there are a lot of overlap between the confidence intervals between females and males for all education groups, total activities is not statistically significant difference between the genders across these education categories.
* Both females and males tend to have a smaller variation of total activities in ages within the 'More than high school' education category compared to the other two education categories.

### Inspection activity over the course of the day
three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex:
```{r}
dat_3|>
  group_by(minute, sex, education)|>
  summarise(mims_value=mean(mims_value))|>
  ggplot(aes(x = minute, y = mims_value, color = sex)) + 
  geom_point(alpha = .75)+
  geom_smooth()+
  facet_grid(.~education)
```

Comments:

* Overall, both females and males tend to have higher MIMS values from 8:30 to 20:00 (i.e. 510~1200 minute) in all education groups.
* MIMS values will increase dramatically since 4:00, and reach the peak at around 8:30. Then it will decrease slightly from 8:30 to 20:00, and drop dramatically after 20:00.
* Both females and males tend to have higher peak MIMS values  within the 'Less than high school' education category compared to the other two education categories. This suggests that educational attainment may be associated with differences in peak MIMS values, with lower education levels potentially correlating with higher peak MIMS scores.
* For "High school equivalent" and "More than high school" groups, females tend to have higher MIMS values than males from 8:30 to 20:00 while MIMS values are similar between genders in 'Less than high school' education category.
